---
title: "final tent"
author: "Cary Ni"
date: "2022-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## Comprehensive dataset
https://sleepdata.org/datasets

# How much sleep for different groups
Here’s a helpful guide from the Centers for Disease Control and Prevention (CDC) 

# National and world insufficient sleep statistics
https://www.cdc.gov/sleep/data-and-statistics/adults.html

https://www.usa.philips.com/c-dam/b2c/master/experience/smartsleep/world-sleep-day/2021/philips-world-sleep-day-2021-report.pdf

# A sleep disorder that causes sleep deprivation
*Insomnia, Parasomnias

# Social relationships
*https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5627640/

*https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3947242/

# Monitoring sleep patterns of healthy people 
*https://physionet.org/content/mmash/1.0.0/


Section 1
Presenting national data (maps/cities)

Section 2
Diseases correlated with insufficient sleep

Section 3
Social and econ factors related to insufficient sleep


# Intended final project
* A report analyzing the prevalence of insufficient sleep in relation to socioeconomic and physiological factors, and an interactive panel to predict the risk of insufficient sleep for individuals

```{r}
timeline_df = readxl::read_excel("Brain.xlsx", range = "G3:I13")
write_csv(timeline_df, "timeline.csv")
```

